Automated topic naming to support analysis of software maintenance activities

### 2010-08-13 icse/icse2011.tex
updated structure. There is no longer a separation between observations and results. More room for discussion now.
Signed-off-by: Neil Ernst <neil@neilernst.net>

 \documentclass[]{sig-alternate}
 \usepackage{booktabs}
% \usepackage[colorinlistoftodos]{todonotes}
 \usepackage{color}

@@

\title{What's in a name? Automated topic naming of software maintenance activities}

\maketitle
\thispagestyle{empty}

\begin{abstract}
%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
% Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
% Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
% However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
% This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
% Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

This paper begins by introducing labelled topic extraction, demonstrating its usefulness compared to other approaches. We show that these labels with their topics can be learned and used to classify other data-sets. We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.

% \section{Background}

% We provide a brief overview of software repository mining and information retrieval.
% This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.

% \subsection{Definitions}
% We will use the following terminology in the rest of the paper.
% A \emph{message} is a block of text written by a developer. In this
% paper, messages will be the CVS and BitKeeper commit log comments made
% when a developer commits changes to files in a repository. A \emph{word
%   distribution} is the summary of a message by word count. Each word
% distribution will be over the words in all messages. However, most
% words will not appear in each message. A word distribution is effectively
% a word count divided by the message size. A \emph{topic} is a word
% distribution, i.e., a set of words that form a word distribution that is
% unique and independent within the set of documents in our total
% corpus. One could think of a topic as a distribution of the centroid
% of a group of messages. In this paper we often summarize topics by the
% top ten most frequent words of their word distribution.  A \emph{trend}
% is one or more similar topics that recur over time.  Trends are
% particularly important, as they indicate long-lived and recurring
% topics that may provide key insights into the development of the
% project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% % area under the ROC 
% % check
% \emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

% \shrink
% \subsection{Topic and Concept Extraction}
% Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
% independent word distributions (topics) from
% 	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
% \begin{comment}
% Ideally these extracted topics 
% correlate with actual development topics that the developers discussed
% during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
% \end{comment}

% Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents. Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear. We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

% In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

% One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.

% \shrink
% \subsection{Supervised learning}
% While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

\section{Methodology}

% To gather the data, we first extract the commit log comments from a  project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows (periods). Each window is subject to LDA analysis. We arbitrarily create 20 topics, vectors of important words according to LDA.

%Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.

Figure \ref{fig:process} gives an outline of the methodology followed.

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{process-model}
 \caption{Research methodology process view.}
  \label{fig:process}
\end{figure*}

To evaluate our approach, we sought candidate systems that were
mature projects and had openly accessible source control repositories.
We also decided to select systems from the same application
domain, as we felt the functional requirements would probably be
broadly similar.
%We wanted to compare two mature software projects, that had open %repositories, within the same application domain. 
% This allows us to partially control for differences in functional %requirements. 
We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}.
%How many developers?
%There are approximately ... developers working on it. 
The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.

\subsection{Generating the data}
For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.

From that data-set, we created an XML file which separated commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation, e.g. \cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. Furthermore, more than $20$ topics quickly became infeasible for inspection.

\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words. We used several different word lists for comparison.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
%There is some debate about the significance and importance of the %terms in this model. 
%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p. %58]{Boegh2008},'' which is sufficient for the purposes of this %research. 
% NOTE I HAD TO CHANGE THE O HERE IN THE BIBTEX
ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.
However, the terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. We therefore took the words from the taxonomy and expanded them.

%Cite mockus et al

To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy. We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets. We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a mailing list about a long-lived web browser project. For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality. For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
%exp2

For the third -- \textsf{exp3} -- list of quality labels, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches. Similarity in WordNet means siblings in a hypernym tree. We do not include these words here for space considerations (but see the Appendix for our data repository). It is not clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.

\begin{table*}
	\centering
\begin{tabular}{c|p{9cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{Qualities and associated signifiers -- WordNet version (\textsf{exp2})}
	\label{tbl:wnsig}

@@

\subsection{Supervised learning}

\label{sec:suplearn}

In order to validate how effective these word-bag approaches to topic labelling would be we had to make a data set to test against.  For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same quality labels as \textsf{exp2} (software qualities). We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period. We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic. For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. We compared against this data-set, but we also used it for our supervised learning based topic classification.
%inter-rater reliability

We first compared our previous analysis using label matching to our manual classifications to get an error rate for that process described below in Section \ref{sec:unsuplabelling}.

For supervised learning, we used a suite of supervised classifiers from WEKA~\cite{weka09}. WEKA contains a suite of machine learning tools such as support vector machines and Bayes nets. We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.

To assess the performance of the supervised learners, we did a 10-fold cross-validation. This is when a set is partitioned into 10 partitions and then each partition is used once as a test set and 9 other times as part of the training set of 9 partitions. We have reported these results below in Section \ref{sec:suplabelling}.

Finally, using this data, we evaluated two research questions (see
Section \ref{sec:comparison}):
\begin{enumerate}
\item Do label frequencies change over time? That is, is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?  
\end{enumerate}

\section{Observations and evaluation}

\subsection{Word list similarity}

% word list similarity

In general, this approach did not work out well as common labels
dominated the less common labels. The related words for
\emph{correctness}, for example, tended to be too lengthy and
non-specific. 
Table \ref{tbl:wordlist} lists results. 
A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

%\begin{comment}
%\XXX{this table should have mysql}
\begin{table}
	\centering
\begin{tabular}{c|c|c|c}
\toprule
Measure &		 \textsf{exp1} &	 \textsf{exp2} &	 \textsf{exp3} \\
\midrule
Topics &			500		 &			500  	 & 		500  \\
Named topics &		281      &			125      &		328  \\
Unnamed topics &	139      &			295      &      92   \\
\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB 7.500}
	\label{tbl:wordlist}

\end{table}
%\end{comment}
%This is the counts

@@

\subsection{Analysis of the unsupervised labelling}

@@

Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we tried to assess whether the manual tag matched the unsupervised label assigned. Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. In general results are poor. Total ROC was slightly better than 0.5, and usability seems relatively accurate.

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{roc-unsup-pretty}
 \caption{ROC values for each NFR, per project and wordlist.}
  \label{fig:maxdb-unsup-results}
\end{figure*}

% Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2. We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.

% \begin{table*}%[ht]
% 	\centering

% \begin{tabular}{c|c|c|c|c|c|c}
% \toprule
% Experiment & Label & F1 & MCC & Precision & Recall & ROC  \\ 
% \midrule
% MaxDB exp2 & portability & 0.228 & 0.182 & 0.520 & 0.146 & 0.553 \\ 
%  & efficiency & 0.217 & 0.125 & 0.237 & 0.200 & 0.558\\ 
%  & reliability & 0.380 & 0.340 & 0.246 & 0.829 & 0.765 \\ 
%  & functionality & 0.095 & 0.083 & 0.250 & 0.059 & 0.521 \\ 
%  & maintainability & 0.092 & 0.123 & 0.571 & 0.050 & 0.520 \\ 
%  & usability & 0.175 & 0.138 & 0.113 & 0.389 & 0.620 \\ 
%  & total & 0.236 & 0.127 & 0.248 & 0.225 & 0.561 \\ 
% \midrule
% MySQL exp2 & portability & 0.138 & 0.211 & 1.000 & 0.074 & 0.537 \\ 
%  & efficiency & 0.345 & 0.327 & 0.476 & 0.270 & 0.625 \\ 
%  & reliability & 0.425 & 0.287 & 0.348 & 0.545 & 0.669 \\ 
%  & functionality & 0.025 & 0.006 & 0.571 & 0.013 & 0.501 \\ 
%  & maintainability & 0.000 & 0.000 & 0.000 & 0.000 & 0.500 \\ 
%  & usability & 0.175 & 0.135 & 0.200 & 0.156 & 0.560 \\ 
%  & total & 0.167 & 0.095 & 0.403 & 0.105 & 0.527 \\ 
% \midrule
% MaxDB exp3 & portability & 0.472 & 0.286 & 0.402 & 0.573 & 0.660 \\ 
%  & efficiency & 0.223 & 0.068 & 0.130 & 0.778 & 0.549 \\ 
%  & reliability & 0.257 & 0.196 & 0.149 & 0.927 & 0.652 \\ 
%  & functionality & 0.236 & 0.187 & 0.138 & 0.824 & 0.665 \\ 
%  & maintainability & 0.338 & 0.112 & 0.266 & 0.463 & 0.566 \\ 
%  & usability & 0.108 & 0.094 & 0.057 & 0.944 & 0.595 \\ 
%  & total & 0.258 & 0.093 & 0.160 & 0.671 & 0.568 \\ 
% \midrule
% MySQL exp3 & portability & 0.413 & 0.170 & 0.564 & 0.325 & 0.574 \\ 
%  & efficiency & 0.158 & 0.105 & 0.089 & 0.703 & 0.608 \\ 
%  & reliability & 0.388 & 0.240 & 0.260 & 0.758 & 0.660 \\ 
%  & functionality & 0.652 & 0.240 & 0.655 & 0.649 & 0.620 \\ 
%  & maintainability & 0.203 & 0.007 & 0.240 & 0.175 & 0.503 \\ 
%  & usability & 0.105 & 0.013 & 0.057 & 0.688 & 0.513 \\ 
%  & total & 0.362 & 0.076 & 0.284 & 0.499 & 0.544 \\
% \bottomrule
% \end{tabular}
% 	\caption{Results for automatic topic labelling. F1 = f-measure, MCC = Matthew's correlation coeff., ROC = Area under ROC curve}
% 	\label{tbl:maxdb-unsup-results}
% \end{table*}
Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. MySQL had reasonable results within \textsf{exp2} for reliability and efficiency. MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve. If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.

%\subsection{Classifying topics}

\subsection{Analysis of the supervised labelling}
\label{sec:suplabelling}

>>>>

 \documentclass[]{sig-alternate}
 \usepackage{booktabs}
\usepackage{subfigure}
 \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes}
 \usepackage{color}

@@

%\title{What's in a name? Automated topic naming of software maintenance activities}
\title{Automated topic naming to support analysis of software maintenance activities}
\maketitle
\thispagestyle{empty}

\begin{abstract}
  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
 Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
 Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
 However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
 This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
 We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
 Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

This paper begins by introducing labelled topic extraction. We show that these labels with their topics can be learned and used to classify other data-sets. We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.

\section{Study design and execution}
Figure \ref{fig:process} gives an outline of the methodology followed. We began by gathering source data; then applied unsupervised and supervised learning techniques to identify topics. We then use these topics to analyse the role of non-functional requirements in software maintenance.

\begin{figure}
  \centering
 \includegraphics[width=.45\textwidth]{process-model}
 \caption{Research methodology process view.}
  \label{fig:process}
\end{figure}

\subsection{Generating the data}
To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. We selected systems from the same application domain, so the functional requirements would be broadly similar. We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code. 
\todo[inline]{why 3.23 and not newer?}

For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. This is the most readily accessible source of project interactions for outside researchers; bug reports and email messages were not available for both projects. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}. Stemming was performed in the later stages of our analysis.

From that data-set, we created an XML file which separated commits into 30 day periods. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. For this set of periods, we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross-cutting collections of words relevant to one or more of our commit messages. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
We set the number of topics to generate to $20$, because past experimentation, e.g. \cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. Furthermore, more than $20$ topics quickly became infeasible for inspection.

\subsection{Generating word lists}
Topics are word distributions: lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. We sought to automate the process of naming the topics.

Accordingly, we tried to associate each topic with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words (a topic might match several keywords). We used several different word lists for comparison, which we call \textsf{exp1,exp2,exp3} in the text which follows.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). There is some debate about the significance and importance of the terms in this model. However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research. The terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. We therefore took the words from the taxonomy and expanded them.

To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms. For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.

For the third list of quality labels,  \textsf{exp3}, we extended the list from \textsf{exp2} using unfiltered WordNet similarity matches. Similarity in WordNet means siblings in a hypernym tree. We do not include these words here for space considerations (but see the Appendix for our data repository). It is not clear the words associated with our labels are specific enough, however: for example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.

\begin{table*}
	\centering
\begin{tabular}{c|p{9cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{NFRs and associated signifiers -- \textsf{exp2}}
	\label{tbl:wnsig}

@@

\subsection{Unsupervised labelling}
% word list similarity
Using our three word lists, we labeled our topics where there was a match between the list and the topic. In general, this approach did not work out well as common labels dominated the less common labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. 
Table \ref{tbl:wordlist} lists broad labeling results for MaxDB as an example. A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

\todo[inline]{this table should have mysql}
\begin{table}
	\centering
\begin{tabular}{c|c|c|c}
\toprule
Measure & \textsf{exp1} & \textsf{exp2} & \textsf{exp3} \\
\midrule
Named topics   & 281 & 125 & 328  \\
Unnamed topics & 139 & 295 & 92   \\
\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB 7.500}
	\label{tbl:wordlist}

\end{table}

@@

\subsubsection{Analysis of the unsupervised labelling}

@@

Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we tried to assess whether the manual tag matched the unsupervised label assigned. Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. Our error is measured using the ROC area-under-curve value. This value is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
  \label{fig:maxdb-unsup-results}
\end{figure*}

Based on these results we find that reliability and usability worked well for MaxDB in \textsf{exp2} and better in \textsf{exp3}. MySQL had reasonable results within \textsf{exp2} for reliability and efficiency. MySQL's results for efficiency did not improve in \textsf{exp3} but other qualities such as functionality did improve. If a \emph{C} grade performance has a ROC value of $0.6$ then most of these tests scored a grade of \emph{C} or less, but our results were still better than random chance.

\subsection{Supervised labelling}
\label{sec:suplearn}
Supervised labelling requires expert analysis of the correct word to assign to a topic. %In order to validate how effective these word-bag approaches to topic labelling would be we created a validation data set. 
 For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). We looked at each period's topics, and assessed what the data -- consisting of the frequency-weighted word lists and messages -- suggested was the topic for that period. We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to this topic. For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. We compared against this data-set, but we also used it for our supervised learning based topic classification. A topic may have more than one matching keyword.
\todo[inline]{inter-rater reliability }

We used a suite of supervised classifiers from WEKA~\cite{weka09}. WEKA is a suite of machine learning tools such as support vector machines and Bayes nets. We also used the multi-labelling add-on for WEKA, Mulan~\cite{mulan}\footnote{\url{http://mlkd.csd.auth.gr/multilabel.html}}. Traditional classifiers map our topics to a single class, whereas Mulan allows for a mixture of classes per topic, which maps to what we observed while manually labelling topics.

To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine learners. The original data is partitioned randomly into ten subsamples. One subsample is used as a training set and evaluated on the other nine subsamples. This is repeated nine more times, with each subsample used once as the training set. We have reported these results below in Section \ref{sec:suplabelling}.

\subsubsection{Analysis of the supervised labelling}
\label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label.  We considered the best learner for a label to be the one which had the highest ROC value for that label. Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning. ZeroR naively chooses the largest category all of the time. It occasionally outperforms non-naive learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. This is why we use ROC values, as they can better represent performance on labels which are not applicable to the majority of samples.

Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have quite different results, as the ROC values for reliability and functionality seem swapped between projects. For both projects Bayesian techniques did the best out of a wide variety of machine learners tested. Discriminative Multinomial Naive Bayes (DMNBtext), Naive Bayes (NaiveBayes) and Multinomial Naive Bayes (NaiveBayesMultinomal) are all based on Bayes's theorem and all assume, naively, that the features are independent. The features we used are word counts per message. One beneficial aspect of this result is that it suggests we can have very fast training and classifying  since Naive Bayes can be calculated in $O(N)$ for $N$ features.
%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

The less-frequently occurring a label the harder it is to get accurate results, due to the high noise level. Nevertheless, these results are better than our previous word bag results of \textsf{exp2} and \textsf{exp3}, because the ROC values are sufficiently higher in most cases (other than MaxDB reliability and MySQL efficiency). The limitation of the approach we took here is that we assume labels are independent; however, labels could be correlated with each other. We also did not evaluate how well the learners performed together.
